Modelling the emergence of whisker barrels

### 2020-07-08 paper/briefpaper/brief.tex
some typos

\cite{karbowski_model_2004} developed a reaction-diffusion style model of how
extrinsic signalling gradients can constrain the emergence of distinct fields
from intrinsic cortical dynamics. Their model defines how the density of
connections $c(x,t)$ and axon branches $a(x,t)$ interact at time $t$, along a
1D anterior-posterior axis $x$, for $N$ thalamocortical projections indexed by
$i$. The model was derived from the assumption that the rates at which $a_i$
and $c_i$ grow are reciprocally coupled. Extending the original 1D model to
simulate arealization on a 2D cortical sheet, we use $a_i(\mb{x},t)$ and
$c_i(\mb{x},t)$, and model synaptogenesis as
%
\begin{equation} \label{eq:dc}
\frac{\partial c_i}{\partial t} =-\alpha c_i +\beta  \left(1 - \sum_{j=1}^{N} c_{j}\right)[a_i]^k.
\end{equation}
%
Accordingly, where the total density of synaptic connections sums to one,
connections decay at rate $\alpha$. Otherwise the connection density increases
non-linearly ($k>1$) with the density of axon branching. Axon branching is
modelled as
%
\begin{equation} \label{eq:da}
\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x}) \MPtwo{+ \chi_i}\right) - \frac{\partial c_i}{\partial t}.
\end{equation}
%
The first term on the right describes the divergence (indicated by
$\nabla\vcdot$) of the quantity in parentheses, which is referred to as the
`flux' of axonal branching. The flux represents diffusion across the cortical
sheet, at rate $D$, and the influence of $M$ molecular signalling fields,
$\rho(\mb{x})$. The influence of a given field (indexed by $j$) on a given
thalamic projection (indexed by $i$), is determined by $\gamma_{i,j}$, which
may be positive or negative in order that axons may branch in the direction of
either higher or lower concentrations. Note that computing the divergence in
simulation requires cells on the cortical sheet to communicate with immedately
adjacent cells only (see \emph{Materials \& Methods}). The second term on the
right represents the coupling between axon branching and synaptogenesis. Here
  $\chi_i=0$ is a placeholder.

@@

\MPthreePred{To further investigate the interplay of genetic and intrinsic
  mechanisms we simulated two well known experimental manipulations of 
barrel development. First, we simulated a seminal barrel duplication
  paradigm}
\citep{shimogori_fibroblast_2005,assimacopoulos_fibroblast_2012}
\MPthreePred{in which the growth factor Fgf8, which is normally expressed at
  the anterior end of the cortical subplate from around E9.5}
\citep{crossley_mouse_1995}, \MPthreePred{is ectopically expressed (by
  electroporation) also at the posterior pole. We assume that
  this results in a mirror of the primary barrel cortex boundary along the
  rostrocaudal axis}
\cite{assimacopoulos_fibroblast_2012} \MPthreePred{and a the mirroring of the
  anterior-posterior guidance gradient $\rho_1$ at the border between them (Fig.\,\ref{fig:fgf8}A). 
  The result after 30000 iterations, and otherwise using the parameters of the example simulation, was two mirror-symmetrical
  barrel fields comprising $2N$ barrels (Fig.\,\ref{fig:fgf8}B), consistent
  with the outcome of the original experiments.}

>>>>

\cite{karbowski_model_2004} developed a reaction-diffusion style model of how
extrinsic signalling gradients can constrain the emergence of distinct fields
from intrinsic cortical dynamics. Their model defines how the density of
connections $c(x,t)$ and axon branches $a(x,t)$ interact at time $t$, along a
1D anterior-posterior axis $x$, for $N$ thalamocortical projections indexed by
$i$. The model was derived from the assumption that the rates at which $a_i$
and $c_i$ grow are reciprocally coupled. Extending the original 1D model to
simulate arealisation on a 2D cortical sheet, we use $a_i(\mb{x},t)$ and
$c_i(\mb{x},t)$, and model synaptogenesis as
%
\begin{equation} \label{eq:dc}
\frac{\partial c_i}{\partial t} =-\alpha c_i +\beta  \left(1 - \sum_{j=1}^{N} c_{j}\right)[a_i]^k.
\end{equation}
%
Accordingly, where the total density of synaptic connections sums to one,
connections decay at rate $\alpha$. Otherwise the connection density increases
non-linearly ($k>1$) with the density of axon branching. Axon branching is
modelled as
%
\begin{equation} \label{eq:da}
\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x}) \MPtwo{+ \chi_i}\right) - \frac{\partial c_i}{\partial t}.
\end{equation}
%
The first term on the right describes the divergence (indicated by
$\nabla\vcdot$) of the quantity in parentheses, which is referred to as the
`flux' of axonal branching. The flux represents diffusion across the cortical
sheet, at rate $D$, and the influence of $M$ molecular signalling fields,
$\rho(\mb{x})$. The influence of a given field (indexed by $j$) on a given
thalamic projection (indexed by $i$), is determined by $\gamma_{i,j}$, which
may be positive or negative in order that axons may branch in the direction of
either higher or lower concentrations. Note that computing the divergence in
simulation requires cells on the cortical sheet to communicate with immedately
adjacent cells only (see \emph{Materials \& Methods}). The second term on the
right represents the coupling between axon branching and synaptogenesis. Here
  $\chi_i=0$ is a placeholder.

@@

\MPthreePred{To further investigate the interplay of genetic and intrinsic
  mechanisms we simulated two well known experimental manipulations of 
barrel development. First, we simulated a seminal barrel duplication
  paradigm}
\citep{shimogori_fibroblast_2005,assimacopoulos_fibroblast_2012}
\MPthreePred{in which the growth factor Fgf8, which is normally expressed at
  the anterior end of the cortical subplate from around E9.5}
\citep{crossley_mouse_1995}, \MPthreePred{is ectopically expressed (by
  electroporation) also at the posterior pole. We assume that
  this results in a mirror of the primary barrel cortex boundary along the
  rostrocaudal axis}
\cite{assimacopoulos_fibroblast_2012} \MPthreePred{and a mirroring of the
  anterior-posterior guidance gradient $\rho_1$ at the border between them (Fig.\,\ref{fig:fgf8}A). 
  The result after 30000 iterations, and otherwise using the parameters of the example simulation, was two mirror-symmetrical
  barrel fields comprising $2N$ barrels (Fig.\,\ref{fig:fgf8}B), consistent
  with the outcome of the original experiments.}
